---
title: "Visualising the RiX Database"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---
This file documents summary statistics of the RiX database, and can be used to indicate data gaps in RiX. Firstly, do we want to focus on one specific country? If not, leave as NA:
```{r}
# iso<-NA
iso<-"SSD"
```
Let's check out some basics on the country specified:
```{r}
source("./GeneralFunctions.R")
# First load the administrative boundaries:
# SSD<-readRDS("../Data/SSD_ADM.Rdata")
# pop<-as(raster::raster("../Data/ssd_ppp_2020_1km_Aggregated_UNadj.tif"),"SpatialPixelsDataFrame")

# Aggregate population to admin 1 boundary level
# SSD$ADM1@data$population<-Grid2ADM(pop,SSD$ADM1)
# Aggregate population to admin 2 boundary level
# SSD$ADM2@data$population<-Grid2ADM(pop,SSD$ADM2)

# adm1er<-SSD$ADM1
# adm1er@data<-dplyr::select(adm1er@data,population)

# list_GDACS<-GetGDACS_API(syear = 2016,fyear = 2022)

# dfGDACS<-FilterGDACS(list_GDACS = list_GDACS,iso3c = "SSD")

st_layers("~/Downloads/hzd-ssd-mh.gpkg")

hazd <- st_read("~/Downloads/hzd-ssd-mh.gpkg",
                    layer = "lss_ssd_fl_adm2")

hazd%<>%dplyr::select(adm1_pcode,
                      adm1_en,
                      adm2_pcode,
                      adm2_en,
                      pop_2019,
                      under18,
                      healthsite_count,
                      fl_pop_rp5,
                      fl_pop_rp20,
                      fl_pop_rp100,
                      fl_pop_rp250,
                      fl_buildings_rp100)
# hazd$GrossNatInc%<>%exp()

SHDI<-xlsx::read.xlsx("~/Downloads/SHDI-Database_SSD.xlsx",sheetName = "SSD")
SHDI%<>%filter(year==max(SHDI$year) & level=="Subnat")%>%dplyr::select(c(4,8:ncol(SHDI)-1))
colnames(SHDI)[2:ncol(SHDI)]<-c("SHDI","HealthI","LivingStandI","EducationI","LifeExp","logGrossNatInc","ExpectedSchoolYrs","MeanSchoolYrs")
if(length(unique(SHDI$GDLCODE))!=length(unique(hazd$adm1_pcode))) stop("Unequal number of p-codes between SHDI and GFDRR-multihazard Risk")
warning("Assuming that order of p-codes is same between SHDI and GFDRR-multihazard Risk")
SHDI$adm1_pcode<-NA
for(i in 1:length(unique(SHDI$GDLCODE))) {
  SHDI$adm1_pcode[SHDI$GDLCODE==unique(SHDI$GDLCODE)[i]]<-as.character(unique(hazd$adm1_pcode)[i])
}
hazd%<>%merge(SHDI,by="adm1_pcode")
bbox<-unname(st_bbox(hazd$geometry))
e <- as(extent(bbox[c(1,3,2,4)]), 'SpatialPolygons')
crs(e) <- "+proj=longlat +datum=WGS84 +no_defs"

# Agricultural Stress Index
shaz <- st_read("../Data/Hazard/hzd-ssd-dr-asi.gpkg",
                                  layer = "hzd_ssd_dr_asi_30years_75th")
ASI<-data.frame(adm1_pcode=unique(shaz$adm1_pcode),
                ASI=sapply(unique(shaz$adm1_pcode),function(id) unique(shaz$asi[shaz$adm1_pcode==id])))
hazd%<>%merge(ASI,by="adm1_pcode")

# Raster files -> aggregate to admin level 2
hpoly<-as(hazd,"Spatial")
# Extreme Temperature hazard
WBGT<-as(raster::raster("../Data/Hazard/hzd-ssd-eh-WBGT_maxTemp.tif"),"SpatialPixelsDataFrame")
hazd$WBGT<-Grid2ADM(WBGT,hpoly,sumFn = "max")
# Wildfires
WF<-as(raster::raster("../Data/Hazard/hzd-ssd-WF-fwi_veg.tif"),"SpatialPixelsDataFrame")
    # STEAL GRID FOR HIGHER RESOLUTION
tmp<-WBGT; tmp$hzd.ssd.eh.WBGT_maxTemp<-Interp2GRID(WBGT,WF); WF<-tmp;rm(tmp)
hazd$WF<-Grid2ADM(WF,hpoly,sumFn = "mean")
# Earthquakes
EQ<-as(crop(raster::raster("../Data/Hazard/SSD_EQ_GAR2017.tif"),e),"SpatialPixelsDataFrame")
hazd$EQ<-Grid2ADM(EQ,hpoly,sumFn = "mean")

st_write(hazd,
         dsn="../Data/SSD_Final",
         layer = "map",
         driver = "ESRI Shapefile")




mad_map <- get_stamenmap(bbox,source = "stamen",zoom=6) # ,maptype = "terrain-background"
q<-ggmap(mad_map,base_layer = ggplot(hazd))+ xlab("Longitude") + ylab("Latitude")


dbase<-xlsx::read.xlsx("~/Downloads/Dataset_Validation.xlsx",sheetName = "Official_RiX_DIIF_XData_Inv_0")
choices<-xlsx::read.xlsx("~/Downloads/RiX-DIIF.xlsx",sheetName = "choices")

CCAfreq<-data.frame(table(unlist(str_split(dbase$Please.select.the.different.themes.that.this.dataset.s..could.be.used.to.analyse,","))))
names(CCAfreq)=c("CodeName","Freq")
CCA<-cbind(arrange(choices[which(choices$name%in%unique(unlist(str_split(dbase$Please.select.the.different.themes.that.this.dataset.s..could.be.used.to.analyse,",")))),c(1:3,7)],name),CCAfreq)
CCAtemplate<-xlsx::read.xlsx("../Data/CompiledFreqTables.xlsx",sheetName = "Sheet1")

SDGfreq<-data.frame(table(unlist(str_split(unlist(str_split(dbase$fin_sdg,",")),";"))))
names(SDGfreq)=c("name","Freq")
xlsx::write.xlsx(merge(choices[,1:3],SDGfreq,by="name"),append = T,
                 "../Data/FreqTables.xlsx", sheetName = "SDG")

JIAFfreq<-data.frame(table(unlist(str_split(unlist(str_split(dbase$fin_jiaf,",")),";"))))
names(JIAFfreq)=c("name","Freq")
xlsx::write.xlsx(merge(choices[,1:3],JIAFfreq,by="name"),append = T,
                 "../Data/FreqTables.xlsx", sheetName = "JIAF")


# Correct names
names(dbase)[6]<-"org_name"
dbase$org_name[dbase$org_name=="World Bank "]<-"World Bank"
dbase$org_name[dbase$org_name=="NASA Socioeconomic Data and Appplications Center"]<-"NASA Socioeconomic Data and Applications Center"
dbase$org_name[dbase$org_name=="IDMC"]<-"Internal Displacement Monitoring Centre (IDMC)"
dbase$org_name[dbase$org_name=="World Intergrated Trade Solution "]<-"World Intergrated Trade Solution"
dbase$org_name[dbase$org_name=="World Trade Intergrated Solution"]<-"World Intergrated Trade Solution"
dbase$org_name[dbase$org_name=="World Trade Intergrated Solution "]<-"World Intergrated Trade Solution"
dbase$org_name[dbase$org_name=="World Intergrated Trade Solutions"]<-"World Intergrated Trade Solution"
dbase$org_name[dbase$org_name=="Famine and Early Warning Systems Network (FEWSNET)"]<-"Famine Early Warning Systems Network (FEWSNET)"
dbase$org_name[dbase$org_name=="GEM Foundation (OpenQuake)"]<-"Global Earthquake Model (GEM)"
dbase$org_name[dbase$org_name=="Global Earthquake Model"]<-"Global Earthquake Model (GEM)"
dbase$org_name[dbase$org_name=="Global Facility for Disaster Reduction and Recovery "]<-"Global Facility for Disaster Reduction and Recovery (GFDRR)"
dbase$org_name[dbase$org_name=="NASA"]<-"National Aeronautics and Space Administration (NASA)"
dbase$org_name[dbase$org_name==" National Aeronautics and Space Administration (NASA)"]<-"National Aeronautics and Space Administration (NASA)"
dbase$org_name[dbase$org_name=="John Hopkins University & Medicine"]<-"John Hopkins Center for Security Health"
dbase$org_name[dbase$org_name=="European Commission "]<-"European Commission"
dbase$org_name[dbase$org_name=="European Commission Copernicus Programme"]<-"European Commission"
dbase$org_name[dbase$org_name=="Global Forest Watch - World Resources Institute (WRI)"]<-"World Resources Institute (WRI)"
dbase$org_name[dbase$org_name=="Global Forest Watch"]<-"World Resources Institute (WRI)"
dbase$org_name[dbase$org_name=="IMF"]<-"International Monetary Fund (IMF)"
org_tot<-data.frame(table(dbase$org_name))%>%arrange(desc(Freq))
xlsx::write.xlsx(org_tot,"../Data/OrgNameFreqTables.xlsx", sheetName = "Sheet1")


# dfGDACS<-data.frame()
#   for (i in 1:length(list_GDACS)){
#     tmper<-list_GDACS[[i]]
#     if(tmper$properties$iso3!=iso) next
#     dfGDACS<-rb
#   }
# leaflet(hazd) %>% addTiles() %>%
#   addPolygons(data=hazd,fillColor =hazd$pop_2019)%>%
#     # opacity = 1.0, fillOpacity = 0.5,
#     # fillColor = ~colorQuantile("YlOrRd", ROMNAM)(ROMNAM),
#     # highlightOptions = highlightOptions(color = "white", weight = 2,
#       # bringToFront = TRUE) %>%
#   setView(31.3, 6.88, zoom = 5) %>%
#   addPopups(
#     31.3, 6.88,
#     'Total Population (2020) = 11,193,729 <br>
#     Life Expectancy at Birth = 58 <br>
#     GDP per capita (USD-2022) = $1,119.7 <br>
#     Unemployment = 13.9%'
#   )
# width & height of world is 256.2^zoomlevel

```
Plot data that provides a country overview, by administrative region. Firstly, total population and population under 18:
```{r}
p1<-q+geom_sf(data=hazd,aes(fill=pop_2019),alpha=0.5)+scale_fill_viridis_c()+labs(fill = "Tot. Pop.")
p2<-q+geom_sf(data=hazd,aes(fill=under18),alpha=0.5)+scale_fill_viridis_c()+labs(fill = "No. Child. (<18)")

gridExtra::grid.arrange(p1,p2,nrow=1)
```
Look at the population exposed to flood hazard frequency, for different flood return periods:
```{r}
p1<-q+geom_sf(data=hazd,aes(fill=fl_pop_rp5),alpha=0.5)+scale_fill_viridis_c()+labs(fill = "Pop. Exp. 5yr Flood")
p2<-q+geom_sf(data=hazd,aes(fill=fl_pop_rp20),alpha=0.5)+scale_fill_viridis_c()+labs(fill = "Pop. Exp. 20yr Flood")
p3<-q+geom_sf(data=hazd,aes(fill=fl_pop_rp100),alpha=0.5)+scale_fill_viridis_c()+labs(fill = "Pop. Exp. 100yr Flood")
p4<-q+geom_sf(data=hazd,aes(fill=fl_pop_rp250),alpha=0.5)+scale_fill_viridis_c()+labs(fill = "Pop. Exp. 250yr Flood")

gridExtra::grid.arrange(p1,p2,p3,p4,nrow=2)
```
Plot the sub-national Human Development Index (HDI), along with it's constituent indices:
```{r}
p1<-q+geom_sf(data=hazd,aes(fill=SHDI),alpha=0.5)+scale_fill_viridis_c()+labs(fill = "HDI")
p2<-q+geom_sf(data=hazd,aes(fill=HealthI),alpha=0.5)+scale_fill_viridis_c()+labs(fill = "Health Ind.")
p3<-q+geom_sf(data=hazd,aes(fill=LivingStandI),alpha=0.5)+scale_fill_viridis_c()+labs(fill = "Living Standard Ind.")
p4<-q+geom_sf(data=hazd,aes(fill=EducationI),alpha=0.5)+scale_fill_viridis_c()+labs(fill = "Education Ind")

gridExtra::grid.arrange(p1,p2,p3,p4,nrow=2)
```
Finally, raw information that was used to build the HDI:
```{r}
p1<-q+geom_sf(data=hazd,aes(fill=LifeExp),alpha=0.5)+scale_fill_viridis_c()+labs(fill = "Life Exp.")
p2<-q+geom_sf(data=hazd,aes(fill=logGrossNatInc),alpha=0.5)+scale_fill_viridis_c()+labs(fill = "GNI (log)")
p3<-q+geom_sf(data=hazd,aes(fill=MeanSchoolYrs),alpha=0.5)+scale_fill_viridis_c()+labs(fill = "Exp. No. School Yrs")
p4<-q+geom_sf(data=hazd,aes(fill=healthsite_count),alpha=0.5)+scale_fill_viridis_c()+labs(fill = "Num. Healthsites")

gridExtra::grid.arrange(p1,p2,p3,p4,nrow=2)
```

Then load and wrangle the Survey123 database:
```{r, include=FALSE, echo=FALSE, message=FALSE}
rix<-xlsx::read.xlsx("~/Downloads/S123_c1aacc6dde7b4b4b8d43607c1417f283_EXCEL(5).xlsx",
                     sheetName = "Official_RiX_DIIF_XData_Inv_0",as.data.frame = T)

rix%<>%dplyr::select(ObjectID,
                     user,
                     In.English..please.provide.the.name.of.the.dataset,
                     des_org_type,
                     Please.provide.the.name.of.the.organisation.that.produced.the.dataset,
                     Please.provide.the.website.URL.of.the.dataset,
                     des_formattype_dataset,
                     des_formatextension_dataset,
                     Please.provide.the.year.when.the.data.was.published..please.also.provide.the.month.if.known.,
                     Please.select.the.different.themes.that.this.dataset.s..could.be.used.to.analyse,
                     des_cca_themes,
                     cca_group,
                     des_cca_group,
                     des_continent_spatial,
                     concat_countries_spatial,
                     des_final_countries_spatial,
                     final_label_countries,
                     des_resolution_spatial,
                     des_res_spat,
                     What.is.the.starting.year.of.the.data.,
                     What.is.the.end.year.of.the.data..Please.also.provide.the.month.if.known.,
                     Which.of.the.following.hazard.type.s..are.present.in.the.data.,
                     Please.select.hazard.cluster.s..present.in.the.data,
                     Please.select.specific.hazard.s..present.in.the.data,
                     des_exposuretype,
                     des_exposuresubtype,
                     des_vultype,
                     des_vulsubtype,
                     des_cctype,
                     des_ccsubtype,
                     des_imptype,
                     des_impsubtype,
                     fin_sdg,
                     des_fin_sdg,
                     fin_jiaf,
                     des_fin_jiaf,
                     EditDate,
                     category,
                     formduration)

names(rix)<-c("id",
              "user",
              "dataset_name",
              "org_type",
              "org",
              "url",
              "format_type",
              "format_ext",
              "pub_yr",
              "cca_tags",
              "cca_themes",
              "cca_group_tag",
              "cca_groups",
              "continent",
              "iso3c",
              "countries",
              "country_label",
              "spat_res_type",
              "spat_res",
              "start_yr",
              "end_yr",
              "haz_type",
              "haz_subtype",
              "haz_subsubtype",
              "exp_type",
              "exp_subtype",
              "vul_type",
              "vul_subtype",
              "cc_type",
              "cc_subtype",
              "imp_type",
              "imp_subtype",
              "sdg_tag",
              "sdg",
              "jiaf_tag",
              "jiaf",
              "submit_date",
              "category",
              "form_duration"
              )

# rix$pub_yr%<>%openxlsx::convertToDate()
# rix$start_yr%<>%openxlsx::convertToDate()

rix%<>%lapply(function(x) {if (is.factor(x)) x<-as.character(x) ; return(x)})%>%as.data.frame(stringsAsFactors=F)
```
Do we need to filter for one country only?
```{r}
if(!is.na(iso)) rix%<>%filter(iso3c==iso)
```


Let's have a look! Firstly, let's have a look at the organisation types:
```{r}

ggplot(rix)+geom_bar(aes(org_type,fill=org_type)) + ggtitle("RiX Organisation Type Bar Chart") +
  scale_fill_discrete(name="Organisation Type")+xlab("Organisation Type")+ylab("Count")+
  theme(axis.text.x=element_blank(),
        plot.title=element_text(hjust=0.5))

```
Now let's look at the different organisation names that we are using the most:
```{r}
tmp<-rix%>%group_by(org)%>%summarise(Frequency=length(org))%>%arrange(desc(Frequency))
colnames(tmp)[1]<-"Organisation"
tmp$Rank<-1:nrow(tmp)
tmp[,c(3,1,2)]
rm(tmp)
```

How about the year of publication?
```{r}
ggplot(rix)+geom_histogram(aes(AsYear(pub_yr)),fill="darkgreen",bins=30) + xlab("Publication Year")
```
What kind of data do we have?
```{r}
ggplot(rix)+geom_bar(aes(histosplitter(spat_res_type,T),fill=histosplitter(spat_res_type,T)))+ ggtitle("Data Types Bar Chart") +
  scale_fill_discrete(name="Dataset/Information Format")+
  theme(axis.text.x=element_blank(),
        plot.title=element_text(hjust=0.5))
```


For the datasets that we are collecting, which data country coverage levels are the most frequent?
```{r}
ggplot(rix)+geom_bar(aes(country_label,fill=country_label))+ ggtitle("RiX Continents Bar Chart") +
  scale_fill_discrete(name="Dataset Coverage")+
  theme(axis.text.x=element_blank(),
        plot.title=element_text(hjust=0.5))
```
For the datasets that we are collecting, are they global, regional, multiple or single countries?
```{r}
ggplot(rix)+geom_bar(aes(country_label,fill=country_label))+ ggtitle("RiX Country Coverage Groups Bar Chart") +
  scale_fill_discrete(name="Coverage Type")+
  theme(axis.text.x=element_blank(),
        plot.title=element_text(hjust=0.5))
```
How about the categories that the datasets span?
```{r}
ggplot(rix)+geom_bar(aes(category,fill=category)) + ggtitle("RiX Data Categories Bar Chart") +
  scale_fill_discrete(name="Category")+
  theme(axis.text.x=element_blank(),
        plot.title=element_text(hjust=0.5))
```
How about the 13 SDG risk areas?
```{r}
ggplot(rix)+geom_bar(aes(sdg_tag,fill=sdg_tag)) + ggtitle("RiX 13 SDG Risk Areas Bar Chart") +
  scale_fill_discrete(name="SDG Risk Area")+
  theme(axis.text.x=element_blank(),
        plot.title=element_text(hjust=0.5))
```
How about the JIAF themes?
```{r}
ggplot(rix)+geom_bar(aes(jiaf,fill=jiaf)) + ggtitle("RiX JIAF Themes Bar Chart") +
  scale_fill_discrete(name="JIAF Themes")+
  theme(axis.text.x=element_blank(),
        plot.title=element_text(hjust=0.5))
```

How long does it take to complete the form (in minutes), including having to resubmit?
```{r}
ggplot(filter(rix,form_duration>0))+geom_histogram(aes(form_duration),fill="darkorange3")+
  scale_x_log10() + xlab("Minutes") + ylab("Count") + ggtitle("Duration to Complete DIIF")+
  theme(plot.title=element_text(hjust=0.5))
```
Here we explore some commonly available data covering South Sudan. Let's look first at EM-DAT

```{r}
emdat<-xlsx::read.xlsx("~/Downloads/emdat_public_2022_06_08_query_uid-1nPHdz.xlsx",
                     sheetName = "emdat data",as.data.frame = T, startRow = 7)
```
Now let's plot the different hazards experienced in South Sudan
```{r}
ggplot(emdat)+geom_bar(aes(Disaster.Type,fill=Disaster.Group))
```
How about number of events in time?
```{r}
ggplot(emdat)+geom_histogram(aes(Start.Year),fill="darkred")
```
How about the number of deaths, per hazard event, per year?
```{r}
ggplot(emdat)+geom_point(aes(Start.Year,as.numeric(Total.Deaths),colour=Disaster.Type,size=as.numeric(Total.Deaths)))
```
How about a boxplot of the number of deaths and homeless per disaster type?
```{r}
ggplot(emdat)+geom_boxplot(aes(Disaster.Type,as.numeric(Total.Deaths),fill=Disaster.Group))
ggplot(emdat)+geom_boxplot(aes(Disaster.Type,as.numeric(No.Affected),fill=Disaster.Group))
```


<!-- Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*. -->

<!-- When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file). -->

<!-- The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed. -->